Automated Dynamic Slice Backsolving

### 2014-03-22 docs/paper/fsuthesis/thesis-template/paper.tex
paper work and idea for strcpy

\title{Automated Dynamic Slice Backsolving}
%\title{A Survey of Fuzzing and New Technique for Dynamic Analysis}
\author{Clark Wood}
\college{College of Arts and Science}
\department{Department of Computer Science} % Delete if no department
\manuscripttype{Thesis}              % [Thesis, Dissertation, Treatise]
\degree{Masters of Science}                 
\semester{Spring}                            % [Spring, Summer, Fall]
\degreeyear{2014}
\defensedate{XXXXXXX}

@@

\committeeperson{Zhi Wang}{Professor Directing Thesis}
%\committeeperson{Beauxeau D'Claune}{Committee Member}
%\committeeperson{Arlip Zarseeld}{Committee Member}

@@

\begin{acknowledgments}
SFS
% Dr. Wang
% Bret Whissel
% Salwan
\end{acknowledgments}

@@

Finding software bugs is a difficult but lucrative endeavour. Fuzzing is one technique for finding bugs, particularly exploitable vulnerabilities, employed by both software companies and nation-state actors. However, when no source code is available, fuzzing complex programs effectively can prove difficult, and as techniques become more sophisticated they require more resources and time. We present a technique, automated dynamic slice backsolving, which, to the best of our knowledge, represents a novel way to reason about not only the location of user input bytes which contribute to interesting program behavior like branching, but also the exact byte value necessary to trigger a desire branch condition. We have implemented this technique in a fuzzer, which is able to reason about large and complex user input and reliably drive program execution down new code paths to explore new portions of compiled code.

%TODO add RE protocol to abstract and intro once implemented

@@

In order to fuzz more smartly, researchers employ several techniques, two of which are significant for this paper. Dynamic Taint Analysis, or DTA, provides valuable information for guided fuzzing. DTA is a form of dynamic analysis, which marks user input as tainted at some level of granularity, and follows this taint as it propagates through the binary. Certain, task-dependent sinks are identified, and if taint flows to one of these sinks, a trace of program execution allows analysis to follow how taint sources are linked to these sinks. Symbolic execution first runs a program concretely to gather constraints on a particular element, then executes the program symbolically, giving these constraints over to some kind of solver which decides what values the element should given the recorded conditions. Symbolic execution improves code coverage, helping to guide fuzzers into deeper, more interesting parts of a program. DTA and symbolic execution, however, are computationally expensive, which leaves room for refinement of techniques by using the technique more wisely or modifying it. Our technique uses DTA to hopefully achieve greater code coverage than random testing, without having to convert code to an intermediate representation or outsource constraints to a solver like symbolic execution.

@@

We have implemented this technique in a proof of concept fuzzer, stubble, which operates on x86 binaries. Stubble dynamically instruments binaries, and automatically generates test files which exercise new paths in the input program. Program input, output, and interesting conditions like segfaults are recorded during execution, and taint propagation is recorded from user input sources to branch sinks. When a tainted branch is found, stubble automically backsolves the execution slice and generates a new test case which will drive the program down the opposite path explored in the current execution. Stubble is able to reason about large amounts of bytes of user input, for instance by performing oracle attacks on binaries which check user input against large secrets, requiring O(n) test cases to solve the secret with no previous information, where n is the length of the secret. Stubble has also been parallized at the process level, allowing for an arbitrary number of independent bytes to be reasoned about concurrently. We believe automated dynamic slice backsolving is a reasonable complementary technique to dynamic taint analysis and symbolic execution for binary fuzzing.

@@

\section{Challenges}

@@

%TODO Limitations - one-way functions like bitwise shifts and checksums, cryptographic hashes. encryption.

@@

@@

>>>>

\title{Automated Dynamic Slice Backsolving}
%\title{A Survey of Fuzzing and New Technique for Dynamic Analysis}
\author{Clark Wood}
\college{College of Arts and Science}
\department{Department of Computer Science} % Delete if no department
\manuscripttype{Thesis}              % [Thesis, Dissertation, Treatise]
\degree{Masters of Science}                 
\semester{Spring}                            % [Spring, Summer, Fall]
\degreeyear{2014}
\defensedate{April 19, 2014}

@@

\committeeperson{Zhi Wang}{Professor Directing Thesis}
%\committeeperson{Xiuwen Liu}{Committee Member}
%\committeeperson{Robert van Engelen}{Committee Member}

@@

\begin{acknowledgments}
SFS
%Zhi Wang
%Bret Whissel
% Salwan
\end{acknowledgments}

@@

Finding software bugs is a difficult but lucrative endeavour, and fuzzing is one technique often employed for finding bugs, particularly exploitable vulnerabilities. However, when no source code is available, fuzzing complex programs effectively can prove difficult, and as techniques become more sophisticated they require more resources and time. We present a technique, automated dynamic slice backsolving, which, to the best of our knowledge, represents a novel way to improve fuzzers' code coverage, without resorting to heavyweight techniques like symbolic execution. We demonstrate this technique with a proof of concept fuzzer which explores paths in a binary to effectively reveal a secret, performing an oracle attack without an oracle. Our approach requires no source code, no manual instrumentation, and no symbolic execution, and, as our results show, can reliably discover and explore new execution paths in a timely fashion. We believe dynamic slice backsolving represents a potentially strong complementary technique for modern fuzzers to implement.

@@

In order to fuzz more smartly, researchers employ several techniques, two of which are significant for this paper. Dynamic Taint Analysis, or DTA, provides valuable information for guided fuzzing. DTA is a form of dynamic analysis, which marks user input as tainted at some level of granularity, and follows this taint as it propagates through the binary. Certain, task-dependent sinks are identified, and if taint flows to one of these sinks, a trace of program execution allows analysis to follow how taint sources are linked to these sinks. Symbolic execution first runs a program concretely to gather constraints on a particular element, then executes the program symbolically, giving these constraints over to some kind of solver which decides what values the element should given the recorded conditions. Symbolic execution improves code coverage, helping to guide fuzzers into deeper, more interesting parts of a program. Symbolic execution, however, is computationally expensive, with naive implementations running in time exponential to the number of branches [C], which leaves room for refinement by using the technique more wisely or modifying it. Our technique uses DTA to hopefully achieve greater code coverage than random testing, without having to convert code to an intermediate representation or outsource constraints to a solver like symbolic execution.

@@

We have implemented this technique in a proof of concept fuzzer, stubble, which operates on x86 binaries. Stubble dynamically instruments binaries, and automatically generates test files which exercise new paths in the input program. Program input, output, and interesting conditions like segfaults are recorded during execution, and taint propagation is recorded from user input sources to branch sinks. When a tainted branch is found, stubble automically backsolves the execution slice and generates a new test case which will drive the program down the opposite path explored in the current execution.

To test stubble's ability to dynamically discover new branches and generate test cases which will drive future runs down these new branches, we have designed several binaries which receive user input in the form of a file. These binaries open and read the file, comparing a subset of the file contents against a secret. If the file contains the secret, then a win condition is printed, otherwise a lose condition is printed. For the proof of concept, this win condition could be thought of as a segfault, or other interesting condition that suggests an exploitable condition exists. We show that stubble is able to discover all paths in the binary, particularly the interesting paths involving the win condition, without any prior knowledge of the secret. Stubble works similarly to an oracle attack, able to discover one character within the secret per run, and preserve work across runs to crack the secret in O(n) tests, where n is the secret length in characters. In contrast, random byte fuzzers, which exhaustively test all user input bytes which their model declares valid, run indefinitely and have no guarantee of discovering the win condition. We have also parallelized stubble at the process level, allowing for an arbitrary number of independent bytes to be reasoned about concurrently. We believe automated dynamic slice backsolving is a reasonable complementary technique to dynamic taint analysis and symbolic execution for binary fuzzing, and that stubble provides a valuable proof of concept for future research into smart fuzzers.

@@

\section{Challenges and Limitations}

@@

As a proof of concept, stubble lacks the featurefulness of a complete fuzzer. In particular, understanding more diverse types of user input, such as command line arguments, would prove valuable.

Note that dynamic slice backsolving requires transformations on tainted values to be easily reversible. Because of various cryptographic properties, encryption and hashing algorithms are not possible to backsolve. In fact, any lossy or not one-to-one operations prove difficult. As an example, the UDP checksum algorithm [AP] calculates the checksum of a packet by iterating over a pseudo-header, summing over each octets 16-bit 1's complement. Because of the repeating summing, an overflow is possible, and any overflow bits are added in later on. This means two non-identical packet headers could result in the same checksum, and so although we can backsolve to the original packet, we cannot make reasonable mutations which will result in the checksum we desire.

@@

% New types of user input like command line arguments. Can get at argv by passing it directly to driver?

@@

	\item [AP] Braden, R., Borman, D., and C. Partridge, "Computing the Internet checksum", RFC 1071, September 1988.
